---
title: "Getting started with pcatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with pcatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 10, fig.height = 8)
library(pcatR)
```

## Read-before-use boundary

The pCAT has 14 items and two response components per item. `pcatR` preserves
those components and derives item-level descriptive categories. It does not
calculate an overall validated pCAT scale score. Open the complete technical
guide with `pcat_user_guide()` before using operational or research data.

## Inspect the instrument and mappings

```{r}
pcat_items("both")
pcat_construct_map("2022", include_secondary = TRUE)
pcat_response_options()
```

## Run the standard workflow

```{r}
dat <- pcat_example_data()
analysis <- pcat_analyse(
  dat,
  group_vars = c("site_id", "timepoint"),
  require_complete = TRUE,
  validation_action = "none"
)
analysis
```

Always inspect validation findings, even when validation does not raise a
condition.

```{r}
pcat_validation_issues(analysis$validation)
head(analysis$summary)
head(analysis$consensus)
```

## Visualize item patterns

```{r, fig.height=10, fig.alt="Diverging pCAT item profiles comparing barrier and facilitator response shares at planning and mid-implementation."}
plot_pcat_profile(
  analysis$classified,
  group_vars = "timepoint",
  label = "cfir_original_construct"
)
```

## Compare repeated assessments

```{r}
change <- pcat_change(
  analysis$classified,
  from = "planning",
  to = "mid_implementation"
)
head(change)
```

## Export reproducible results

```{r, eval=FALSE}
pcat_write_analysis(
  analysis,
  path = "pcat_analysis_outputs",
  overwrite = TRUE
)
```

The export includes an analysis manifest, validation findings, classified
responses, item summaries, consensus diagnostics, an action-plan worksheet,
and an optional multi-page profile PDF.
