---
title: "Interpretation, privacy, and reporting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interpretation, privacy, and reporting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(pcatR)
```

## Item-level interpretation

`pcatR` classifies complete responses as strong barrier, weak barrier, neutral,
weak facilitator, or strong facilitator. The `pcat_display_code` from -2 to +2
supports plotting and within-item transition descriptions only. Do not sum it
across items or describe it as a validated scale.

## Agreement and polarization

```{r}
classified <- pcat_classify(pcat_example_data())
summary <- pcat_summarise(
  classified,
  group_vars = c("site_id", "timepoint")
)
consensus <- pcat_consensus(summary)
head(consensus[c(
  "site_id", "timepoint", "item_id", "dominant_side",
  "agreement_share", "polarized", "normalized_entropy", "consensus_label"
)])
```

Consensus labels are descriptive decision aids. They depend on user-visible
thresholds and should be interpreted with counts and response distributions.

## Longitudinal comparison

```{r}
change <- pcat_change(
  classified,
  from = "planning",
  to = "mid_implementation"
)
head(change[c(
  "respondent_id", "item_id", "from_pcat_class5", "to_pcat_class5",
  "delta_display_code", "transition"
)])
```

These paired transitions describe ordered category movement. They are not
causal effects and the display-code differences are not interval-scale change
scores.

## Privacy and suppression

Use coded identifiers and review free-text comments for sensitive content.
Respondent heatmaps can reveal individual patterns and should not be shared for
small or identifiable teams. Summary suppression is available:

```{r}
suppressed <- pcat_summarise(
  classified,
  group_vars = c("site_id", "timepoint"),
  suppress_below = 5
)
head(suppressed)
```

Suppression affects the returned summary table only. It does not transform or
de-identify source data, classified responses, heatmaps, or profile PDFs.

## Reporting checklist

Report the implementation effort assessed, assessment timing, respondent roles,
valid denominators, missing effects, grouping variables, consensus thresholds,
suppression rule, mapping version, and software version. Describe action-plan
strategies as locally reviewed candidates rather than algorithmically selected
recommendations.
